### 2018-10-11 paper/complexity.tex
checked new material

In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test \fixme{In
\cite{DBLP:conf/icfem/CavalcantiG07}, we work with minimal acceptances. I
didn't check \cite{Hennessy:1988:ATP:50497}.} \fxnote{jp: we have studied
\cite{DBLP:conf/icfem/CavalcantiG07} now in more detail -- you are right, of
course, but there is a confusion of terms: you are also using the minimal
hitting sets we handle here in our joint paper, but you used the term
``minimal acceptance'' for them. This is not in line with the definition of
minimal acceptances as presented in \cite{Hennessy:1988:ATP:50497} and in
Roscoe's books. Here, in our joint paper, minimal acceptances are introduced
as defined in Roscoe's books. I have changed the text accordingly.} {\it
every} non-empty subset of $\Sigma$ whose events cannot be completely refused
in a given process state of the reference model; this leads to a worst-case
estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
offered to the SUT in the last step of the test execution, so our approach
reduces the number of test executions in comparison
to~\cite{Hennessy:1988:ATP:50497} by a factor of $1/{\sqrt{2\pi
\card{\Sigma}}}$. In Fig.~\ref{fig:minhita}, the reduction is visualised by a
function plot of $2^{\card{\Sigma}}$ versus
$\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.

>>>>

In~\cite{Hennessy:1988:ATP:50497}, it is suggested to test {\it every}
non-empty subset of $\Sigma$ whose events cannot be completely refused in a
given process state of the reference model; this leads to a worst-case
estimate of $2^{\card{\Sigma}}-1$ for the number of different sets to be
offered to the SUT in the last step of the test execution, so our approach
reduces the number of test executions in comparison
to~\cite{Hennessy:1988:ATP:50497} by a factor of $1/{\sqrt{2\pi
\card{\Sigma}}}$. In Fig.~\ref{fig:minhita}, the reduction is visualised by a
function plot of $2^{\card{\Sigma}}$ versus
$\frac{2^\card{\Sigma}}{\sqrt{2\pi \card{\Sigma}}}$.

### 2018-10-11 paper/finitecomplete-failures.tex
checked new material

\paragraph{Step~2.} For proving that a test execution can never deadlock without a 
$\epass$ event, it has to be shown that a test execution can neither block
at branch (\ref{eq:ufc}) nor at branch (\ref{eq:ufd}).

\paragraph{Step~2.1} Suppose that the test execution blocks at branch (\ref{eq:ufc})
after having run through trace $s$ with $\#s < j$. 
Since  $P\lessdet_T Q$ by assumption, $s$ is a trace of $P$, thus 
$U_F(j)/s = U_F(j,\#s,G(P)/s)$ according to Lemma~\ref{lemma:ufproperties}. Therefore,
$U_F(j)/s$  can enter  branch (\ref{eq:ufc}) with any event from $[G(P)/s]^0$.
Since we assume that $(Q\parallel[\Sigma] U_F(j))/s$ deadlocks, this means that
$[G(P)/s]^0$ is {\it not} a hitting set of $\minaccs(Q/s)$, because otherwise at least
on $e\in [G(P)/s]^0$ would be accepted by $Q/s$. Now suppose that $\minhits(G(P)/s) = \varnothing$. Then   branch (\ref{eq:ufb}) can be entered, and the test 
stops after $\epass$. Otherwise, if $\minhits(G(P)/s) \neq \varnothing$, let
$H\in\minhits(G(P)/s)$. Since $H$ only contains elements that are contained in some minimal acceptance of $P/s$, and all these minimal acceptances are subsets of 
$[G(P)/s]^0$, $H$ is a subset of $[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, 
is not a hitting set of $\minaccs(Q/s)$, also $H$ is not a hitting set of 
$\minaccs(Q/s)$.
Now this is a contradiction to Lemma~\ref{lemma:hseta}, since $Q\ conf\ P$ by assumption,
so $H$ should also be a (not necessarily minimal) hitting set in $\minaccs(Q/s)$.
This proves that the test execution cannot block  at branch (\ref{eq:ufc})
without being able to pass the test by entering branch (\ref{eq:ufb}).

@@

Note that the line of reasoning in this proof requires that $Q$ is free of
livelocks, because otherwise a $\epass$ event might not become visible,
due to unbounded sequences of hidden events performed by $Q$. 
\xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
exhaustive for the fault model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$. These
cases can be characterised as follows:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of these cases that at least one test execution
of some $(Q\parallel[\Sigma] U_F(j))$ with $j < pq$ ends with the $\efail$
event or deadlocks. We do this by analysing the product graph of the
reference process $P$ and the SUT process $Q$: any trace
$s\in\trc(Q)\cap\trc(P)$ gives rise to a path labelled by the events of $s$
through this product graph. Any error can be detected after running through
such a trace and then either observing an event outside $[P/s]^0$ (this is
the violation described by Case~1) or identifying an illegal acceptance $A_Q$
as specified in Case~2. It is not guaranteed, however, that trace $s$ is
short enough to be executed by one of the test cases $U_F(j)$ with $0\le j <
pq$. So, it has to be shown that for any $s$ leading to an error situation
specified by Case~1 or Case~2, there exists a trace $u$ of maximal length
$pq-1$ leading to the same error.

@@

\medskip
\noindent {\bf Case~2.} We note that trace $s$ is again a trace of the
product graph $G$, but we do not know its length. Again, by applying
Lemma~\ref{lemma:reachproduc}, we know that the state $G/s$ can be reached by
a trace $u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. We consider
the test $U_F(\# u)$, for which $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because
of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$  always performs branch
(\ref{eq:ufc}) until the trace $u$ has been completely processed.  
 $U_F(\#u,\#u,G(P)/u)$ may execute
branches (\ref{eq:ufa}) or (\ref{eq:ufd}) only:~assumption
(\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has at least one
non-empty minimal acceptance. From (\ref{eq:minhitminaccempty}) we know that
this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and
we observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$. As a
consequence, branch (\ref{eq:ufb}) cannot be taken because its guard
condition evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$. The guard condition
$(k < j)$ for branch (\ref{eq:ufc}) evaluates to $\isf$ for
$U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the test
always fails. If branch (\ref{eq:ufd}) is executed, the test   deadlocks and therefore 
fails for the
execution where 
$Q/u$ selects the minimal acceptance $A_Q$ as specified in (\ref{eq:accsnotcontained})
and $U_F(\#u,\#u,G(P)/u)$ selects 
a minimal hitting set $H\in\minhits(P/u)$ which has an empty intersection with   $A_Q$.
The existence of
such an $H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a
consequence,   $(Q\parallel[\Sigma]U_F(\# u))/u$,   cannot produce the
$\epass$-event in this execution; 
this means that the test fails. The complete testing assumption guarantees that 
this execution really occurs if $(Q\parallel[\Sigma]U_F(\# u))$ is executed sufficiently often.
This concludes the proof.
\xbox 
\end{proof}
%
Our notion of tests can be specialised to deal with traces refinement, as we
explain in Section~\ref{sec:finitecomplete}.
The next section presents and example.

>>>>

\paragraph{Step~2.} For proving that a test execution can never deadlock without a
$\epass$ event, it has to be shown that a test execution can neither block at
branch (\ref{eq:ufc}) nor at branch (\ref{eq:ufd}). These cases are
considered separately below.

\paragraph{Step~2.1} Suppose that the test execution blocks at branch (\ref{eq:ufc})
after having run through trace $s$ with $\#s < j$. Since  $P\lessdet_T Q$ by
assumption, $s$ is a trace of $P$, thus $U_F(j)/s = U_F(j,\#s,G(P)/s)$
according to Lemma~\ref{lemma:ufproperties}. Therefore, $U_F(j)/s$  can enter
branch (\ref{eq:ufc}) with any event from $[G(P)/s]^0$. Since we assume that
$(Q\parallel[\Sigma] U_F(j))/s$ deadlocks, this means that $[G(P)/s]^0$ is
{\it not} a hitting set of $\minaccs(Q/s)$, because otherwise at least one
$e\in [G(P)/s]^0$ would be accepted by $Q/s$. Now suppose that
$\minhits(G(P)/s) = \varnothing$. Then   branch (\ref{eq:ufb}) can be
entered, and the test stops after $\epass$. Otherwise, if $\minhits(G(P)/s)
\neq \varnothing$, let $H\in\minhits(G(P)/s)$. Since $H$ contains only
elements that are contained in some minimal acceptance of $P/s$, and all
these minimal acceptances are subsets of $[G(P)/s]^0$, $H$ is a subset of
$[G(P)/s]^0$ as well. Since $[G(P)/s]^0$, however, is not a hitting set of
$\minaccs(Q/s)$, also $H$ is not a hitting set of $\minaccs(Q/s)$. Now this
is a contradiction to Lemma~\ref{lemma:hseta}, since $Q\ conf\ P$ by
assumption, so $H$ should also be a (not necessarily minimal) hitting set in
$\minaccs(Q/s)$. This proves that the test execution cannot block  at branch
(\ref{eq:ufc}) without being able to pass the test by entering branch
(\ref{eq:ufb}).

@@

Note that the line of reasoning in this proof requires that $Q$ is free of
livelocks, because otherwise a $\epass$ event might not become visible,
due to unbounded sequences of hidden events performed by $Q$.
\xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
A test suite $\TS_F$ specified as in Theorem~\ref{th:failurestest} is
exhaustive for the fault model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$. According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$. These
cases can be characterised as follows:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of these cases that at least one test execution
of some $(Q\parallel[\Sigma] U_F(j))$ with $j < pq$ ends with the $\efail$
event or deadlocks. We do this by analysing the product graph of the
reference process $P$ and the SUT process $Q$: any trace
$s\in\trc(Q)\cap\trc(P)$ gives rise to a path labelled by the events of $s$
through this product graph. Any error can be detected after running through
such a trace and then either observing an event outside $[P/s]^0$ (this is
the violation described by Case~1) or identifying an illegal acceptance $A_Q$
(as specified in Case~2). It is not guaranteed, however, that trace $s$ is
short enough to be executed by one of the test cases $U_F(j)$ with $0\le j <
pq$. So, it has to be shown that for any $s$ leading to an error situation
specified by Case~1 or Case~2, there exists a trace $u$ of maximal length
$pq-1$ leading to the same error.

@@

\medskip
\noindent {\bf Case~2.} We note that trace $s$ is again a trace of the
product graph $G$, but we do not know its length. Again, by applying
Lemma~\ref{lemma:reachproduc}, we know that the state $G/s$ can be reached by
a trace $u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. We consider
the test $U_F(\# u)$, for which $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because
of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$  always performs branch
(\ref{eq:ufc}) until the trace $u$ has been completely processed.
 $U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
only:~(\ref{eq:accsnotcontained}) implies that $P/s$ has at least one
non-empty minimal acceptance. By (\ref{eq:minhitminaccempty}) this is
equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and we
observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$. As a
consequence, branch (\ref{eq:ufb}) cannot be taken because its guard
condition evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$. The guard condition
$(k < j)$ for branch (\ref{eq:ufc}) evaluates to $\isf$ for
$U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the test
always fails. If branch (\ref{eq:ufd}) is executed, the test   deadlocks and
therefore fails for the execution where $Q/u$ selects the minimal acceptance
$A_Q$ as specified in (\ref{eq:accsnotcontained}) and $U_F(\#u,\#u,G(P)/u)$
selects a minimal hitting set $H\in\minhits(P/u)$ that has an empty
intersection with $A_Q$. The existence of such an $H$ is guaranteed because
of Lemma~\ref{lemma:hseta}. As a consequence, $(Q\parallel[\Sigma]U_F(\#
u))/u$ cannot produce the $\epass$ event in this execution; this means that
the test fails. The complete testing assumption guarantees that this
execution really occurs if $(Q\parallel[\Sigma]U_F(\# u))$ is executed
sufficiently often. This concludes the proof. \xbox
\end{proof}
%
Our notion of tests can be specialised to deal with traces refinement, as we
explain in Section~\ref{sec:finitecomplete}.
The next section presents and example.

### 2018-10-11 paper/preliminaries.tex
checked new material

% -----------------------------------------------------------------------------
\subsubsection*{Minimal Hitting Sets of Minimal Acceptances}
In this article, we are interested in the minimal hitting sets of minimal acceptances;
for these, the abbreviated notation
\[
\minhits(P/s) = \minhits(\minaccs(P/s))
\]
is used.

The minimal hitting sets of minimal acceptances may be alternatively
characterised by means of the failures of a process. To this end,
define for any collection ${\cal C}\subseteq 2^\Sigma$ of subsets from
$\Sigma$
\begin{equation}\label{eq:minsub}
\minsub({\cal C}) = \{ A\in {\cal C}~|~\forall B\in {\cal C}:
B\subseteq A\Rightarrow B = A \}.
\end{equation}
This definition has originally been in~\cite{DBLP:conf/icfem/CavalcantiG07}:
the collection 
$\minsub({\cal C})$ contains all those sets of the collection ${\cal C}$ that are not true subsets of other members of ${\cal C}$. With this definition, the relation
between failures and minimal hitting sets of minimal acceptances is established in the following lemma.
%
\begin{lemma}

>>>>

% -----------------------------------------------------------------------------
\subsubsection*{Minimal Hitting Sets of Minimal Acceptances}
In this article, we are interested in the minimal hitting sets of minimal acceptances;
for these, the abbreviated notation
\[
\minhits(P/s) = \minhits(\minaccs(P/s))
\]
is used. The minimal hitting sets of minimal acceptances may be alternatively
characterised by means of the failures of a process as is done
in~\cite{DBLP:conf/icfem/CavalcantiG07}. To this end,
in~\cite{DBLP:conf/icfem/CavalcantiG07}, the authors define, for any
collection ${\cal C}\subseteq 2^\Sigma$ of subsets from $\Sigma$
\begin{equation}\label{eq:minsub}
\minsub({\cal C}) = \{ A\in {\cal C}~|~\forall B\in {\cal C}:
B\subseteq A\Rightarrow B = A \}.
\end{equation}
The collection $\minsub({\cal C})$ contains all those sets of 
${\cal C}$ that are not true subsets of other members of ${\cal C}$. With
this definition, the relation between failures and minimal hitting sets of
minimal acceptances is established in the following lemma.
%
\begin{lemma}
